---
title: "Attud - Brief Report"
author: "Henrique Pinto Gomide"
date: "08-27-2015"
output:
  word_document: default
  pdf_document:
    toc: yes
  html_document:
    highlight: tango
    number_sections: yes
    theme: spacelab
    toc: yes
---

# Introduction

This is the first report about the survey "International Needs for Treating Tobacco Use". The survey has as the main objective understand the needs of tobacco treatment provider.

Be advised, this is not the final report, so it may contain errors.

# Method

All technology (until this date) are freely available on internet. This report was generated using R. 

# Results
```{r include=FALSE, message=FALSE}
# Load dataframe
library("ggplot2")
library("sjPlot")
attud <- read.csv("attudtestalfa_2015_08_27_12_11_02.csv", na.strings="n/a")
attud <- attud[-113, ] # Remove my test
sjp.setTheme(axis.textsize.y = .9, axis.textsize.x = .8)
```

## Sample characteristics

We have gotten so far `r dim(attud)[1]` participants.

## Age
```{r echo = FALSE, message=FALSE, warning=FALSE}
# Convert negative value into NA attud$group5.group8.q15. I checked the case and it seems that was a real person, but the age was wrong.
attud$group5.group8.q15 <- ifelse(attud$group5.group8.q15 < 0, NA, attud$group5.group8.q15)
ggplot(attud, aes(x=group5.group8.q15)) + geom_histogram(aes(fill = ..count..), binwidth = 6) + xlab("Age (years)") + ylab("Frequency")
```

## Sex

```{r echo = FALSE, message=FALSE, warning=FALSE}
attud$group5.group8.q16 <- factor(attud$group5.group8.q16, labels = c("Male", "Female"))
sjp.frq(attud$group5.group8.q16, geom.colors = "#fc8d59")
```

## What is your profession?

```{r echo = FALSE, message=FALSE, warning=FALSE}
attud$group5.group8.q17 <- factor(attud$group5.group8.q17, labels = c("Physician", "Nurse","Psychologist", "Pharmacist", "Social worker", "Dentist", "Nurse assistant", "Community health worker", "Other"))
sjp.frq(attud$group5.group8.q17, geom.colors = "#91bfdb", coord.flip = TRUE,  sort.frq = "asc")
```

## Are you a member of Attud?

```{r echo = FALSE, message=FALSE, warning=FALSE}
attud$group5.group8.q19 <- factor(attud$group5.group8.q19, labels = c("Yes", "No", "Not sure"))
sjp.frq(attud$group5.group8.q19, geom.colors = "#fc8d59")
```

# Practioner Role and Trainning

## How did you learn how to treat tobacco dependence?

```{r echo = FALSE, message=FALSE, warning=FALSE}
question1  <- attud[, 1:7]
q1labels  <- c("Studied written materials", "Studied information on web", "Received instruction during my health profession education", "Completed an online training course designed for tobacco treatment", "Attended an in-person training designed for tobacco treatment", "Never learned how to treat tobacco", "None of the above")
sjp.likert(question1, catcount = 2, geom.colors = c("#fc8d59","#91bfdb"), legendLabels = c("No","Yes"), axisLabels.y = q1labels, gridBreaksAt = .4)
```

##  What tobacco treatment activities do you currently perform?
```{r echo = FALSE, message=FALSE, warning=FALSE}
question2  <- attud[, 9:14]
q2labels  <- c("Treat tobacco users","Train others how to treat tobacco users", "Administer programs that treat tobacco", "Conduct research on tobacco treatment", "Develop policy related to tobacco treatment", "Other")
sjp.likert(question2, catcount = 2, geom.colors = c("#fc8d59","#91bfdb"), legendLabels = c("No","Yes"), axisLabels.y = q2labels, gridBreaksAt = .4)
```

# Tobacco Treatment Training in Your Country

## Think back to your professional training curriculum. About how many hours of instruction in how to treat tobacco dependence were provided?

```{r echo = FALSE, message=FALSE, warning=FALSE}
attud$group5.q3 <- factor(attud$group5.q3, labels = c("0 hours", "1-2 hours", "3-4 hours", "5-6 hours", "More than 6 hours"))
sjp.frq(attud$group5.q3, geom.colors = "#91bfdb", coord.flip = TRUE, labelPos = "inside")
```

## Outside of health care provider education, is any form of training for treating tobacco dependence available in your country?

```{r echo = FALSE, message=FALSE, warning=FALSE}
attud$group5.q4 <- factor(attud$group5.q4, labels = c("Yes", "No", "Not sure"))
# sjp.frq(attud$group5.q4, geom.colors = "#fc8d59")
```

## Overall, how would you rate access to tobacco treatment training in your country?

```{r echo = FALSE, message=FALSE, warning=FALSE}
# Something may be wrong here. In the questionnaire are 6 options, but in the data just 4.
#attud$group5.q5 <- factor(attud$group5.q5, labels = c("Very Poor", "Poor", "Good", "Very Good"))
#sjp.frq(attud$group5.q5, geom.colors = "#91bfdb", coord.flip = TRUE, labelPos = "inside")
```

## Overall, how would you rate the quality of the tobacco treatment training that is available in your country?

```{r echo = FALSE, message=FALSE, warning=FALSE}
attud$group5.q6 <- factor(attud$group5.q6, labels = c("No access", "Very poor access", "Poor access", "Good access", "Very good access", "Not sure"))
sjp.frq(attud$group5.q6, geom.colors = "#91bfdb", coord.flip = TRUE, labelPos = "inside")
```

## Does your country have any form of professional certification for tobacco treatment?

```{r echo = FALSE, message=FALSE, warning=FALSE}
attud$group5.q8 <- factor(attud$group5.q8, labels = c("Yes", "No", "Not sure"))
sjp.frq(attud$group5.q8, geom.colors = "#fc8d59")
```


# Guidelines for Tobacco Treatment

## Does your country have national guidelines for the treatment of tobacco dependence?

```{r echo = FALSE, message=FALSE, warning=FALSE}
attud$group5.group7.q11 <- factor(attud$group5.group7.q11, labels = c("Yes", "No", "Not sure"))
sjp.frq(attud$group5.group7.q11, geom.colors = "#fc8d59")
```

# Survey Collection

## Access over time

```{r echo = FALSE, message=FALSE, warning=FALSE}
attud$X_submission_time  <- as.character(attud$X_submission_time)
attud$X_submission_time <- gsub('T.*', '', attud$X_submission_time)
attud$time  <- as.Date(attud$X_submission_time, format="%Y-%m-%d")
ggplot(attud, aes(time)) + geom_density(colour = "#91bfdb", fill="#91bfdb") + labs(x = "Month", y = "Frequency") 
```
