---
title: "Attud - Brief Report - Updated"
author: "Henrique Pinto Gomide"
date: "09-01-2015"
output:
  html_document:
    highlight: tango
    number_sections: yes
    theme: spacelab
    toc: yes
  pdf_document:
    toc: yes
  word_document: default
---

# Introduction

This is final report on the survey "International Needs for Treating Tobacco Use". The survey aimed to understand the needs of tobacco treatment provider.

# Methods

All tools used to collect, evaluate quality, analyze data were freeware or/and Open-source. For data collection we have used Ona (https://ona.io/), for data analysis R (R Core Team, 2015) using ggplot2 and sjPlot packages. This report was created using RMarkdown.

# Results

```{r include=FALSE, message=FALSE}

####################################
# Load libraries, open dataframe
####################################

# Load dataframe
require("ggplot2")
require("sjPlot")

# Set theme for charts
sjp.setTheme(axis.textsize.y = .9, axis.textsize.x = .8)

# Set outputs for three decimal places
options(digits = 1)

# Read all data but Spanish
attud <- read.csv("attudRawData.csv", na.strings=c("n/a", "NA"), stringsAsFactors = TRUE)

# Map
world <- read.csv("freqTablebyCountry.csv", na.strings=c("n/a", "NA"), stringsAsFactors = FALSE)

#####################################
# Fix small errors
#####################################

# Convert negative value into NA attud$group5.group8.q15. I checked two cases with negative ages. Although they were inserted wrong they seem valid - #125 and #157.

## Age
attud$group5.group8.q15 <- as.numeric(attud$group5.group8.q15)
attud$group5.group8.q15 <- ifelse(attud$group5.group8.q15 < 0, NA, attud$group5.group8.q15)

## Gender
attud$group5.group8.q16 <- factor(attud$group5.group8.q16, labels = c("Male", "Female"))

## Profession
attud$group5.group8.q17 <- factor(attud$group5.group8.q17, labels = c("Physician", "Nurse","Psychologist", "Pharmacist", "Social worker", "Dentist", "Nurse assistant", "Community health worker", "Other"))

## Map
world <- world[,-1]
world$continent <- world$region
world$continent <- gsub("Eastern|South-Central|Western|Southern|Oceania;|Northern|Central East|West|Southeast|South-East|Middle East,|", "", world$continent)
world$continent <- gsub("Arabian Peninsula, Middle East|Middle East|ern Europe,", "Asia", world$continent)
world$continent <- gsub("North North America", "North America", world$continent)
world$continent <- gsub("North  South America|ern South America", "South America", world$continent)
world$continent <- gsub("Greater Antilles, Caribbean", "Central America", world$continent)
world$continent <- gsub("Australia", "Oceania", world$continent)
world$continent <- gsub("Africa, Asia|Asia  Asia", "Asia", world$continent)
world$continent <- gsub("\\s+","",world$continent)
names <- levels(factor(world$continent))
continent <- data.frame(as.character(names), as.numeric(tapply(world$Freq, world$continent, sum)))
names(continent) <- c("continent", "Freq")
continent <- subset(continent, continent != "Antarctica" ) # Remove an error.
continent$percen <- round(continent$Freq / sum(continent$Freq),3)*100
```


## Participants

The data reported here were collected from `r dim(attud)[1]` valid cases. One case was considered invalid (test submission), thus removed. Most of participants were women `r prop.table(table(attud$group5.group8.q16))[2]*100`% with mean age of `r mean(attud$group5.group8.q15, na.rm = TRUE)` (SD = `r sd(attud$group5.group8.q15, na.rm = TRUE)`). 

## Profession
```{r echo = FALSE, message=FALSE, warning=FALSE}
sjp.frq(attud$group5.group8.q17, geom.colors = "#91bfdb", coord.flip = TRUE,  sort.frq = "asc", axisLimits.y = c(0,100))
```

## Region of the World
```{r echo = FALSE, message=FALSE, warning=FALSE}
#ggplot(continent, aes(x = names, y = Freq)) + geom_bar(stat="identity")
```

## Are you a member of Attud?
```{r echo = FALSE, message=FALSE, warning=FALSE}
attud$group5.group8.q19 <- factor(attud$group5.group8.q19, labels = c("Yes", "No", "Not sure"))
sjp.frq(attud$group5.group8.q19, geom.colors = "#fc8d59")
```

# Practioner Role and Trainning

## How did you learn how to treat tobacco dependence?

```{r echo = FALSE, message=FALSE, warning=FALSE}
question1  <- attud[, 1:7]
q1labels  <- c("Studied written materials", "Studied information on web", "Received instruction during my health profession education", "Completed an online training course designed for tobacco treatment", "Attended an in-person training designed for tobacco treatment", "Never learned how to treat tobacco", "None of the above")
sjp.likert(question1, catcount = 2, geom.colors = c("#fc8d59","#91bfdb"), legendLabels = c("No","Yes"), axisLabels.y = q1labels, gridBreaksAt = .4)
```

##  What tobacco treatment activities do you currently perform?
```{r echo = FALSE, message=FALSE, warning=FALSE}
question2  <- attud[, 9:14]
q2labels  <- c("Treat tobacco users","Train others how to treat tobacco users", "Administer programs that treat tobacco", "Conduct research on tobacco treatment", "Develop policy related to tobacco treatment", "Other")
sjp.likert(question2, catcount = 2, geom.colors = c("#fc8d59","#91bfdb"), legendLabels = c("No","Yes"), axisLabels.y = q2labels, gridBreaksAt = .4)
```

# Tobacco Treatment Training in Your Country

## Think back to your professional training curriculum. About how many hours of instruction in how to treat tobacco dependence were provided?

```{r echo = FALSE, message=FALSE, warning=FALSE}
attud$group5.q3 <- factor(attud$group5.q3, labels = c("0 hours", "1-2 hours", "3-4 hours", "5-6 hours", "More than 6 hours"))
sjp.frq(attud$group5.q3, geom.colors = "#91bfdb", coord.flip = TRUE, labelPos = "inside")
```

## Outside of health care provider education, is any form of training for treating tobacco dependence available in your country?

```{r echo = FALSE, message=FALSE, warning=FALSE}
attud$group5.q4 <- factor(attud$group5.q4, labels = c("Yes", "No", "Not sure"))
# sjp.frq(attud$group5.q4, geom.colors = "#fc8d59")
```

## Overall, how would you rate access to tobacco treatment training in your country?

```{r echo = FALSE, message=FALSE, warning=FALSE}
# Something may be wrong here. In the questionnaire are 6 options, but in the data just 4.
#attud$group5.q5 <- factor(attud$group5.q5, labels = c("Very Poor", "Poor", "Good", "Very Good"))
#sjp.frq(attud$group5.q5, geom.colors = "#91bfdb", coord.flip = TRUE, labelPos = "inside")
```

## Overall, how would you rate the quality of the tobacco treatment training that is available in your country?

```{r echo = FALSE, message=FALSE, warning=FALSE}
attud$group5.q6 <- factor(attud$group5.q6, labels = c("No access", "Very poor access", "Poor access", "Good access", "Very good access", "Not sure"))
sjp.frq(attud$group5.q6, geom.colors = "#91bfdb", coord.flip = TRUE, labelPos = "inside")
```

## Does your country have any form of professional certification for tobacco treatment?

```{r echo = FALSE, message=FALSE, warning=FALSE}
attud$group5.q8 <- factor(attud$group5.q8, labels = c("Yes", "No", "Not sure"))
sjp.frq(attud$group5.q8, geom.colors = "#fc8d59")
```


# Guidelines for Tobacco Treatment

## Does your country have national guidelines for the treatment of tobacco dependence?

```{r echo = FALSE, message=FALSE, warning=FALSE}
attud$group5.group7.q11 <- factor(attud$group5.group7.q11, labels = c("Yes", "No", "Not sure"))
sjp.frq(attud$group5.group7.q11, geom.colors = "#fc8d59")
```

# Survey Collection

## Access over time

```{r echo = FALSE, message=FALSE, warning=FALSE}
attud$X_submission_time  <- as.character(attud$X_submission_time)
attud$X_submission_time <- gsub('T.*', '', attud$X_submission_time)
attud$time  <- as.Date(attud$X_submission_time, format="%Y-%m-%d")
ggplot(attud, aes(time)) + geom_density(colour = "#91bfdb", fill="#91bfdb") + labs(x = "Month", y = "Frequency") 
```

# Countries

```{r}
## Import list with uncoded countries
uncoded  <- read.csv("R/uncodedCountries.csv")
## Import list with coded countries
coded <- read.csv("R/codedCountries.csv", col.names = c("code", "local_name"))
## Merge the two lists
listCountries  <- merge(coded, uncoded, by = "local_name")
## Table results from attud questionnaire
teste  <- as.data.frame(table(attud$group5.group8.q18))
## Pick good names for dataframe
colnames(teste) <- c("code", "Freq")
### Convert vars into character
teste$code  <- as.character(teste$code)
listCountries$code  <- as.character(listCountries$code)
listCountries$local_name  <- as.character(listCountries$local_name)
# Merge table and list of codes
table1  <- merge(teste, listCountries, by = "code")

# Subset the data for coutries that answered the survey at least once 
table2  <- subset(table1, table1$Freq > 0)
# Create data from two variables - frequency and English name
finalCountryTable  <- table2[, c("Freq", "en_name", "region")]
# Sort the data
sorted  <- finalCountryTable[order(-finalCountryTable$Freq),  ]
print(sorted)
# Row 13 is from Antarctica! What to do in this case?
write.csv(sorted, "freqTablebyCountry.csv")
```


